---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ggaxistable

ggaxistable allows for a table to be added on either axis of a ggplot figure. Such tables are commonly used in forest plots and in other places.

## Installation

You can install the released version of ggaxistable from [CRAN](https://CRAN.R-project.org) with:

``` r
library(devtools)
devtools::install_github("nicokist/ggaxistable")
```

## Example

First we create some data.
```{r messages=FALSE, warning=FALSE, }
suppressPackageStartupMessages(library(dplyr))
library(forcats)
library(ggplot2)
library(ggaxistable)

ORs <- tibble(
  Name = as.factor(paste("Group", LETTERS[1:26])),
  `Odds Ratio` = 10^rnorm(n = 26, sd = 0.1),
  Upper = 10^(log10(`Odds Ratio`) + runif(26, min = 0.1)),
  Lower = 10^(log10(`Odds Ratio`) - runif(26, min = 0.1))) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(Name = fct_rev(Name))


```

Then we plot this and add a table on the y axis. Note the use of `axis_table` to create the table and `theme(axis.text.y = element_text(hjust = 0, family = "mono", color = "black"))` to ensure a fixed-width font is used.

```{r forestplot_y_axis, dpi=300}
ggplot(ORs) +
  aes(y = Name, yend = Name) +
  geom_point(aes(x = `Odds Ratio`)) +
  geom_segment(aes(x = Lower, xend = Upper)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10() +
  axis_table(ORs, y = "Name", selected_cols = c("Name", "Odds Ratio")) +
  xlab("Odds Ratio") +
  ylab("") +
  theme_classic() +
  theme(
    axis.text.y = element_text(hjust = 0, family = "mono", color = "black"),
    axis.ticks.y = element_blank()
  )
```

If desired, the table can also be added on the x axis.

```{r forestplot_x_axis, dpi=300}
ggplot(ORs) +
  aes(x = Name, xend = Name) +
  geom_point(aes(y = `Odds Ratio`)) +
  geom_segment(aes(y = Lower, yend = Upper)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10() +
  axis_table(ORs, x = "Name", selected_cols = c("Name", "Odds Ratio")) +
  xlab("") +
  ylab("Odds Ratio") +
  theme_classic() +
  theme(
    axis.text.x = element_text(hjust = 0, family = "mono", color = "black", angle = -90),
    axis.ticks.x = element_blank()
  )
```

